Temperature Measurements in Optical Tweezer Experiments

### 2017-01-14 doc/draft1.tex

    E &=& \sum_{i=1}^N \mathbf{v}_i^2\\
    \lambda &=& \sqrt{\frac{3(N-1)T}{E}}\\
    \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i 

@@

\subsection{eHEX}
The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\DeltaQ$. 
The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\DeltaQ$.\\
Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\

>>>>

 \label{eq:rescale1}   E &=& \sum_{i=1}^N \mathbf{v}_i^2 \\
 \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{E}} \\
 \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  

@@

        \includegraphics[scale=0.3]{images/verlet_output_te.pdf}
        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 
        \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
    jumps are happening.}
        \label{fig:energystart}

\subsection{eHEX}
The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
algorithm. In fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_te.pdf}
        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
            and the values for the energy on the right hand side. }
        \label{fig:ehex_te}
    \end{center}
\end{figure}
The comparison of different values of $\Delta Q$ can be seen in fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
been shown in fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
        \label{fig:ehex_dq}
    \end{center}
\end{figure}

\subsection{Barostat}
The barostat, as described in the previous section, keeps the system from overheating. The algorithm is applied at the same time as the eHEX algorithm
which leads to the fact, that the system has time to heat up, until the particles reach the nano crystal. This can be seen in the temperature
development in fig.~\ref{fig:baro1}.
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_te.pdf}
        \caption{}
        \label{fig:baro1}
    \end{center}
\end{figure}
